Ms 1

### 2017-11-01 ms1/ms1_display.Rmd
refactor display

**swo**: Double-check that number.

There are **swo: PDES** and first fills.

@@

@@

kable(inapprop_changes)
```

@@

drug_trends = read_tsv('tables/trend_models.tsv') %>%
  filter(term=='year') %>%
  arrange(estimate) %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  mutate(ci=1.96*std.error,
         ci_lo=estimate-ci,
         ci_hi=estimate+ci) %>%
  mutate(y=est2pct(estimate),
         ymin=est2pct(ci_lo),
         ymax=est2pct(ci_hi)) %>%
  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax))

drug_trends %>%
  select(abx, y, ymin, ymax) %>%
  mutate_if(is.numeric, function(x) format(x, digits=2, nsmall=2)) %>%
  kable(caption='Drug trends: Percent change from 2011 to 2014')

@@

@@

# ***Earlier content

### All fills

## Methodology

First, I looked at all encounters recorded in the Carrier Line File and the Outpatient file. A large fraction, I think 78%, of PDEs had an encounter 0-7 days before.

Next, I limited by analysis to encounters who HCPCS code indicated an E&M visit, and I grouped diagnostic codes into categories using Fleming-Dutra eTable 2. I determined the trends in the prevalence of these diagnosis categories by counting the diagnoses per beneficiary per year and by using Poisson regressions that account for the unusual covariates.

I analyzed trends in prescribing practice by computing the fraction of encounters in each diagnosis category that were followed by PDEs for each drug. I computed adjusted trends using logistic regression.

Finally, I considered only those PDEs that are downstream of those E&M diagnoses. For each antibiotic, I computed the fraction of PDEs that have each diagnosis category upstream.

To compute appropriateness, I classified each PDE with the maximum appropriateness of all its linked diagnoses. I put PDEs with no associated infectious disease (i.e., FD) diagnoses into their own, lower tier.

## Numbers of diagnoses

The fraction of diagnoses that have E&M HCPCS codes:

```{r hcpcs_count}
encounter_counts = read_tsv('raw_data/dx_cat_count.tsv') %>%
  summarize(n_encounters=sum(n_encounters),
            n_em_encounters=sum(n_encounters[hcpcs_em]),
            f_em_encounters=weighted.mean(hcpcs_em, weights=n_encounters))

hcpcs_count %>% kable
```

This equals a total of `r prettyNum(sum(encounter_counts$n_encounters), big.mark=',')` encounters.

The fraction of first-fill PDEs that have upstream:

```{r pde_with_dx_upstream}
read_tsv('tables/pde_with_dx_upstream.tsv') %>%
  group_by(has_dx, has_em_dx) %>%
  summarize(n=sum(n)) %>%
  ungroup() %>%
  mutate(f=n/sum(n), cumf=cumsum(f), cumf1=1-cumf) %>%
  kable
```

## Common diagnoses by drug

For each drug, I asked what fraction of PDEs are associated with each diagnosis. This gives you a sense that, say, the diagnoses that give rise to azithromycin are bronchitis, URI, etc. The values are percents in 2011. A row can add up to more than 100% because one PDE can be associated with multiple diagnoses.

```{r dx_from_pde}
dx_from_pde = read_tsv('tables/dx_from_pde.tsv') %>% 
  spread(present, n) %>%
  mutate(pct=`TRUE`/(`TRUE` + `FALSE`)*100 %>% round(2)) %>%
  select(antibiotic, dx_cat, pct) %>%
  spread(dx_cat, pct)

dx_from_pde %>% kable
```

## Trends in prescribing practice

For each diagnosis and each drug, I ask what fraction of diagnoses are followed by a PDE for that drug in 2011 and 2014. I compute the percent change in those fractions.

Second, I do logistic regressions. The unit of analysis is the diagnosis. The outcome is whether a PDE for that drug follows a diagnosis in that category. The main predictor is year; covariates are as above. The adjusted percent change is between 2011 and 2014. (I computed this adjusted change, which is a relative risk, using the odds ratio from the regression and the mean probability of PDE for that diagnosis in 2011. This strictly incorrect: the mean relative risk is the mean of the risks computed using the predicted probabilities for each 2011 diagnosis, *not* the relative risk computed using the mean probability in 2011.)

```{r dx_to_pde}
dx_to_pde_unadj = read_tsv('tables/dx_to_pde_table.tsv') %>%
  filter(year %in% c(2011, 2014)) %>%
  select(dx_cat, antibiotic, year, f=f_encounters_with_abx) %>%
  spread(year, f)

dx_to_pde_adj = read_tsv('tables/dx_to_pde_trends.tsv') %>%
  filter(term=='year') %>%
  mult_est(3) %>%
  mutate(ci=1.96*std.error,
         ci_low=estimate-ci,
         ci_high=estimate+ci) %>%
  select(dx_cat, antibiotic, estimate, ci_low, ci_high) %>%
  arrange(dx_cat, antibiotic)

dx_to_pde = inner_join(dx_to_pde_unadj,
                       dx_to_pde_adj,
                       by=c('dx_cat', 'antibiotic')) %>%
  # convert model estimates to odds ratios, then to risk ratios,
  # then to adjusted fractional changes, then to percentages
  mutate(or=exp(estimate),
         or_low=exp(ci_low),
         or_high=exp(ci_high),
         rr=risk_ratio_f(or, `2011`),
         rr_low=risk_ratio_f(or_low, `2011`),
         rr_high=risk_ratio_f(or_high, `2011`),
         adj_change=rr-1,
         adj_change_low=rr_low-1,
         adj_change_high=rr_high-1,
         pct_change=(`2014`-`2011`)/`2011`) %>%
  mutate_if(is.numeric, function(x) round(x*100, 2)) %>%
  mutate(adj_change=sprintf('%.2f (%.2f -- %.2f)', adj_change, adj_change_low, adj_change_high)) %>%
  select(dx_cat, antibiotic, `2011`, `2014`, pct_change, adj_change) %T>%
  kable
```

## Combined tables

```{r}
x = read_tsv('tables/dx_cat_counts.tsv') %>%
  mutate(value=n_dx/n_bene*1000) %>%
  group_by(dx_cat) %>%
  summarize(value=mean(value)) %>%
  # convert to character to fit with other parts
  mutate(value=as.character(round(value))) %>%
  spread(dx_cat, value) %>%
  mutate(antibiotic='dx', key='mean_per_1kyr')

y = dx_from_pde %>%
  mutate(key='dx_from_pde') %>%
  mutate_if(is.numeric, function(x) as.character(round(x, 2)))

z = dx_to_pde %>%
  gather('key', 'value', -dx_cat, -antibiotic) %>%
  spread(dx_cat, value)

combined_table = bind_rows(x, y, z) %>%
  mutate(antibiotic=factor(antibiotic, levels=c('dx', top_abx)),
         key=factor(key, levels=c('mean_per_1kyr', 'dx_from_pde', '2011', '2014', 'pct_change', 'adj_change'))) %>%
  arrange(antibiotic, key)
```

### Respiratory stuff

All respiratory diagnoses fell, bronchitis the most sharply. Azithromycin prescribing fell for all diagnoses, while levofloxacin, amox, and amox/clav prescribing rose for all diagnoses.

```{r respiratory}
combined_table %>%
  filter(antibiotic %in% c('dx', 'azithromycin', 'levofloxacin', 'amoxicillin', 'amoxicillin/clavulanate')) %>%
  select(antibiotic, key, asthma, bronchitis, other_resp, pharyngitis, pneumonia, sinusitis, uri) %>%
  kable
```

### UTIs

UTIs declined somewhat in prevalence. Ciprofloxacin prescribing remained nearly constant, nitrofurantion fell 3%, and Bactrim rose 4-6%.

```{r uti}
combined_table %>%
  filter(antibiotic %in% c('dx', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'nitrofurantoin')) %>%
  select(antibiotic, key, uti, uti_t3) %>%
  kable
```

### SSTI

Although skin and soft tissue infections declined in prevalence, prescribing of clindamycin, cephalexin, Bactrim, amox, and amox/clav increased, with one exception: amox/clav use for Tier 2 SSTIs stay nearly constant. Clindamycin was the biggest winner.

```{r ssti}
# swo: I removed amox and amox/clav because they mostly
# showed trends in Tier 3, which is confusing
combined_table %>%
  filter(antibiotic %in% c('dx', 'clindamycin', 'cephalexin', 'trimethoprim/sulfamethoxazole')) %>%
  select(antibiotic, key, ssti, ssti_t3) %>%
  kable
```

### GI

GI indications became less prevalent, and cipro prescribing fell 1-2%. Only a small amount of cipro is used for GI indications: compare the 5-6% of cipro PDEs preceded by GI indications with the 14-20% preceded by UTIs.

```{r gi}
combined_table %>%
  filter(antibiotic %in% c('dx', 'ciprofloxacin')) %>%
  select(antibiotic, key, gi, gi_t3) %>%
  kable
```

>>>>

```{r pde_counts}
read_tsv('raw_data/firstfill.tsv') %>%
  group_by(firstfill) %>%
  summarize(n_pde=sum(n_pde)) %>%
  #arrange(desc(firstfill)) %>%
  mutate(cum_n=prettyNum(cumsum(n_pde), big.mark=',')) %>%
  kable(caption='PDEs by firstfill (or refill)')
```

@@

drug_trends = read_tsv('tables/trend_models.tsv') %>%
  filter(term=='year') %>%
  arrange(estimate) %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  mutate(ci=1.96*std.error,
         ci_lo=estimate-ci,
         ci_hi=estimate+ci) %>%
  mutate(y=est2pct(estimate),
         ymin=est2pct(ci_lo),
         ymax=est2pct(ci_hi)) %>%
  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax))
```

@@

kable(inapprop_trends, caption='Percent changes in in/appropriate prescribing')
```

@@

```{r drug_trends_display}
drug_trends %>%
  select(abx, y, ymin, ymax) %>%
  mutate_if(is.numeric, function(x) format(x, digits=2, nsmall=2)) %>%
  kable(caption='Drug trends: Percent change from 2011 to 2014')

@@

```{r prescribing_practice}
dx_counts_2011 = read_tsv('raw_data/dx_counts.tsv') %>%
  left_join(select(pop_chars, year, n_bene), by='year') %>%
  mutate(dkpy=round(n_encounter/n_bene*1000)) %>%
  filter(year==2011) %>%
  select(dx_cat, dkpy)

dx_from_pde_2011 = read_tsv('raw_data/dx_from_pde.tsv') %>%
  filter(year==2011) %>%
  mutate(f_with_dx=n_pde_with_encounter/n_pde_with_any_encounter) %>%
  select(antibiotic, dx_cat, year, f_with_dx)

dx_to_pde_table = read_tsv('tables/dx_to_pde_table.tsv') %>%
  select(-n_encounters) %>%
  filter(year %in% c(2011, 2014)) %>%
  gather('abx', 'f_with_abx', azithromycin:clindamycin)

# swo: convert ORs to RRs!
rr2pct = function(x) (x - 1) * 100

dx_to_pde_trends = read_tsv('tables/dx_to_pde_trends.tsv') %>%
  filter(term=='year') %>%
  mutate(estimate=3*estimate,
         std.error=sqrt(3)*std.error) %>%
  mutate(ci=1.96*std.error,
         ci_lo=estimate-ci,
         ci_hi=estimate+ci) %>%
  mutate(or=   exp(estimate),
         or_lo=exp(ci_lo),
         or_hi=exp(ci_hi)) %>%
  left_join(filter(dx_to_pde_table, year==2011), by=c('dx_cat', 'abx')) %>%
  mutate(rr=   risk_ratio_f(or,    f_with_abx),
         rr_lo=risk_ratio_f(or_lo, f_with_abx),
         rr_hi=risk_ratio_f(or_hi, f_with_abx)) %>%
  mutate(y=rr2pct(rr),
         ymin=rr2pct(rr_lo),
         ymax=rr2pct(rr_hi)) %>%
  mutate(change=sprintf('%.2f (%.2f -- %.2f)', y, ymin, ymax)) %>%
  select(dx_cat, abx, change)
```

```{r combined_table}
format2 = function(x) format(round(x, 2), digits=2, nsmall=2)

part1 = dx_counts_2011 %>%
  mutate(dkpy=as.character(dkpy)) %>%
  spread(dx_cat, dkpy) %>%
  mutate(abx='dx', key1='dkpy', key2='')

part2 = dx_from_pde_2011 %>%
  # throw out some rare drugs
  filter(!is.na(dx_cat)) %>%
  mutate(f_with_dx=format2(f_with_dx*100)) %>%
  spread(dx_cat, f_with_dx) %>%
  mutate(key1='pde2dx', key2=as.character(year)) %>% rename(abx=antibiotic)

part3 = dx_to_pde_table %>%
  mutate(f_with_abx=format2(f_with_abx*100)) %>%
  spread(dx_cat, f_with_abx) %>%
  mutate(key1='dx2pde', key2=as.character(year))

part4 = dx_to_pde_trends %>%
  spread(dx_cat, change) %>%
  mutate(key1='dx2pde_trend', key2='')

combined_table = bind_rows(part1, part2, part3, part4) %>%
  mutate(abx=factor(abx, levels=c('dx', abx_levels)),
         key1=factor(key1, levels=c('dkpy', 'pde2dx', 'dx2pde', 'dx2pde_trend'))) %>%
  arrange(abx, key1, key2)
```

## Respiratory

```{r respiratory}
combined_table %>%
  select(abx, key1, key2, pneumonia, sinusitis, uri, bronchitis) %>%
  filter(abx %in% c('dx', 'azithromycin', 'levofloxacin', 'amoxclav')) %>%
  kable
```

@@

## Other respiratory

```{r other_respiratory}
combined_table %>%
  select(abx, key1, key2, pharyngitis, asthma, other_resp) %>%
  filter(abx %in% c('dx', 'azithromycin', 'levofloxacin', 'amoxclav')) %>%
  kable
```

## UTI

```{r uti}
combined_table %>%
  select(abx, key1, key2, uti, uti_t3) %>%
  filter(abx %in% c('dx', 'ciprofloxacin', 'tmpsmx', 'nitrofurantoin')) %>%
  kable
```

## SSTI

```{r ssti}
combined_table %>%
  select(abx, key1, key2, ssti, ssti_t3) %>%
  filter(abx %in% c('dx', 'cephalexin', 'tmpsmx', 'clindamycin')) %>%
  kable
```
